---
title: "cjc_code_sharing"
author: "J Brophy"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
link-citations: yes
linkcolor: red
urlcolor: blue
bibliography: references.bib
csl: vancouver-brackets.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```

## P valve and S curves

The topic of P values and S curves is discussed by [@RN5434]

There is a `R` package, `concurve`, that contains many useful functions and graphs but I have elected to write my own as they can be more easily customized.

## Functions to map P value & S values curves

```{r}
# load standard libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)
library(metafor)
library(meta)
library(epiR)
options(dplyr.summarise.inform=FALSE)

# p value graph function
pv_graph <- function(hr, uci, lci) {
    se <- (log(uci)-log(lci))/(2*1.65)         
    x <- seq(0.01, 0.50,by = .005)
    p1 <- exp(log(hr) - (qnorm(x) * se))
    p2 <- exp(log(hr) + (qnorm(x) * se))
    p <- data.frame(x, p2, p1)
    g <- ggplot(p, aes( p2, x)) +
        geom_line() +
        geom_line(aes(p1, x)) +
        scale_x_continuous(trans='log10') +
        ylab("p value \n one sided") +
        xlab("Hazard ratio (Log scale)") +
        labs (title="P value function") +
        geom_hline(yintercept=c(.005,.025,0.05,0.10), color = "red") +
        annotate("text", x=.7,y=.01, label="99% CI") +
        annotate("text", x=0.74,y=.04, label="95% CI") +
        annotate("text", x=0.78,y=.06, label="90% CI") +
        annotate("text", x=.82,y=.11, label="80% CI") +
        geom_vline(xintercept=1.0, color = "green") +
        annotate("text", x=1,y=.4, label="null hypothesis") +
        theme_bw()
    return(g)
}

# p value graph function
s_graph <- function(hr, uci, lci){
    se <- (log(uci)-log(lci))/(2*1.65)         #.86 0.72 to 1.04 (log se)
    x <- seq(0.01, 0.50,by = .005)
    lci <- exp(log(hr) - (qnorm(x) * se))
    uci <- exp(log(hr) + (qnorm(x) * se))
    lci <- rev(lci)
    hr <- rev(c(uci, lci))
    yy <- 2*x
    yy <- c(yy,rev(yy))
    ss <- -log(yy, base=2)
    df1 <- data.frame(hr,ss)
    df1 <- df1[-297,]
    s <- ggplot(df1, aes( hr,ss)) +
        geom_line() +
        xlim(0.01,1.2) +
        scale_x_continuous(trans='log10') +
        ylab("Bits of information against HR (binary S value)") +
        xlab("Hazard ratio (Log scale)") +
        labs (subtitle = "S-Values (surprisals) for a range of hazard ratios (HR)") +
        geom_vline(xintercept=1.0, color = "green") +
        annotate("text", x=1,y=.4, label="null hypothesis") +
        theme_bw()
        return(s)
}

# function to make S dataset (already intrinsically done in pv_graph and s_graph) but this allows it to be easily exported

s.make <- function(h, uci, lci){
    x <- seq(0.0001, 0.50,length.out = 100)   # p value  x <- seq(0.01, 0.50,by = .005) 
    se <- (log(uci)-log(lci))/(2*1.65) 
    p1 <- exp(log(h) - (qnorm(x) * se))
    p2 <- exp(log(h) + (qnorm(x) * se))
    p2 <- rev(p2)
    hr <- rev(c(p1,p2[-1]))
    prob <- c(2*x, rev(2*x)[-1])
    s <- -log(prob, base=2)
    df1 <- data.frame(hr, prob, s)
    return(df1)
}

# mle function
mle <- function(hr, uci, lci, x){
    d = (log(uci/lci))/(2*(1.96))
    z = log(hr/x)/d
    z_sq <-z*z 
    mle = exp(z_sq/2)
    return(mle)
}

#lr function
lr <- function(hr, uci, lci, x){
    d = (log(uci/lci))/(2*(1.96))
    z = log(hr/x)/d
    z_sq <-z*z 
    return(z_sq)
}

```

## Example - Stich graphical results at 5 years
The 5 year data is available here [@RN5261]

```{r}
# Stich 5 years  #.86 0.72 to 1.04 (log se)

# pv graph
stich_pv_5 <- pv_graph(0.86, 1.04, 0.72) +
    labs (title="Stich trial results 2011", 
          subtitle = "P value function for HR = 0.86, 95% CI 0.72 to 1.04") 

pdf("stich_pv_5.pdf")
stich_pv_5
dev.off()

ggsave("stich_pv_5.png")
stich_pv_5

# S graph
stich_s_5 <- s_graph(0.86, 1.04, 0.72) +
  labs(title="Stich trial results 2011") +
    annotate("text", x=.8,y=1, label="Maximum likelihood estimate (HR=0.86)\n has the least refutational evidence \n against it (0 bits)") +
    geom_segment(aes(x = .86, y = 0.8, xend = .86, yend = 0.015),
               arrow = arrow(length = unit(0.5, "cm")),color="red")

pdf("stich_s_5.pdf")
stich_s_5
dev.off()

ggsave("stich_s_5.png")
stich_s_5
```

## Stich graphical results at 10 years
The 10year data is available here [@RN5263]


```{r fig.width=12, fig.height=12}

# Stich 10 years

# pv graph
stich_pv_10 <- pv_graph(0.84, 0.97, 0.73) +
    labs (title="Stich trial results 2016", 
          subtitle = "P value function for HR = 0.84, 95% CI 0.73 to 0.97") 

pdf("stich_pv_10.pdf")
stich_pv_10
dev.off()

ggsave("stich_pv_10.png")
stich_pv_10

# S graph
stich_s_10 <- s_graph(0.84, 0.97, 0.73) +
  labs(title="Stich trial results 2016") +
    annotate("text", x=.8,y=1, label="Maximum likelihood estimate (HR=0.84)\n has the least refutational evidence \n against it (0 bits)") +
    geom_segment(aes(x = .84, y = 0.8, xend = .84, yend = 0.015),
               arrow = arrow(length = unit(0.5, "cm")),color="red")

pdf("stich_s_10.pdf")
stich_s_10
dev.off()

ggsave("stich_s_10.png")
stich_s_10

# combining graphs
library(patchwork)

stich_comb <- stich_pv_5 + stich_s_5 + stich_pv_10 + stich_s_10 +plot_layout(nrow = 2, byrow = TRUE) +
  plot_annotation(title = 'The surprising story about the Stich trial') + 
  plot_annotation(tag_levels = 'A')

pdf("stich_comb.pdf")
stich_comb
dev.off()

ggsave("stich_comb.png")
stich_comb

```
## Stich Tables

```{r}
stich5 <- s.make(.86, 1.04,.72)
# make table - values extracted from data frame created from s.make
hr <- c(1.1, 1.0, 0.9, 0.86, .75)
pr <- c(0.028, 0.176, 0.68, 1.0, .22)
s <- -log(pr, base=2)
df <- data.frame(HR=hr, P=pr,S=s)
row.names(df) <- (c("10% increase risk", "no association", "10% decrease in risk", "point estimate", "25% decrease in risk"))

df
kable(df, col.names=c("HR","P value (2 sided)","S")) %>%
  kableExtra::kable_styling() %>% 
  add_header_above(c("Stich 5 years" = 4))

z <- qnorm(1-pr/2)
df$lr <- z^2
df$mlr <- exp(df$lr/2)
kable(df, col.names=c("HR","P value (2 sided)","S", "Likelihood ratio", "Maximum likelihood ratio")) %>%
  kableExtra::kable_styling() %>% 
  add_header_above(c("Stich 5 years" = 6))

################
stich10 <- s.make(0.84, 0.97, 0.73)
hr <- c(1.1, 1.0, 0.9, 0.84, .75)
pr <- c( .005, 0.08, 0.84, 1.0, .38)

s <- -log(pr, base=2)
df <- data.frame(HR=hr, P=pr,S=s)
row.names(df) <- (c("10% increase risk", "no association", "10% decrease in risk", "point estimate", "25% decrease in risk"))

df
kable(df, col.names=c("HR","P value (2 sided)","S")) %>%
  kableExtra::kable_styling() %>% 
  add_header_above(c("Stich 10 years" = 4))

z <- qnorm(1-pr/2)
df$lr <- z^2
df$mlr <- exp(df$lr/2)
kable(df, col.names=c("HR","P value (2 sided)","S", "Likelihood ratio", "Maximum likelihood ratio")) %>%
  kableExtra::kable_styling() %>% 
  add_header_above(c("Stich 10 years" = 6))

```

## Fragility index

Consider the PRAMI trial [@RN5276] with a highly significant p value.
This result is not robust as misclassifying only 4 patients eliminates the conventional statistical significance. 

```{r}
# PRAMI example
dat <- matrix(c(11, 27, 223, 204), nrow=2)
rownames(dat) <- c("PCI", "No PCI"); colnames(dat) <- c("Dead", "Alive")
dat
epi.2by2(dat = as.table(dat), method = "cohort.count", 
         conf.level = 0.95, units = 100, outcome = "as.columns")


dat4 <- matrix(c(15, 27, 219, 204), nrow=2)
rownames(dat) <- c("PCI", "No PCI"); colnames(dat) <- c("Dead", "Alive")
dat4
epi.2by2(dat = as.table(dat4), method = "cohort.count", 
         conf.level = 0.95, units = 100, outcome = "as.columns")


dat5 <- matrix(c(16, 27, 218, 204), nrow=2)
rownames(dat) <- c("PCI", "No PCI"); colnames(dat) <- c("Dead", "Alive")
dat5
epi.2by2(dat = as.table(dat5), method = "cohort.count", 
         conf.level = 0.95, units = 100, outcome = "as.columns")

```

## References